### 2019-04-11 paper/derivatives.tex
Changes to eigenvectors

Derivatives of pressure with respect to $\tau$, $\epsilon$, and n. \\
\begin{footnotesize}
\begin{eqnarray}
	\frac{\partial{P}}{\partial{\epsilon}} &=& \left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}\left(\frac{\partial{P}}{\partial{T}}\right) \\
	\frac{\partial{P}}{\partial{n}} &=& \frac{m_B}{\rho} \left[ \frac{\partial{P}}{\partial{y_e}} -
          \frac{\partial{\epsilon}}{\partial{y_e}}\left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}\left(\frac{\partial{P}}{\partial{T}}\right)\right]\\
	\frac{\partial{P}}{\partial{\tau}} &=& -\tau^{-2} \left[ \frac{\partial{P}}{\partial{\rho}} - \frac{y_e}{\rho} \left( \frac{\partial{P}}{\partial{y_e}} -
          \frac{\partial{\epsilon}}{\partial{y_e}} \left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}\left(\frac{\partial{P}}{\partial{T}}\right)\right) -
          \frac{\partial{\epsilon}}{\partial{\rho}}\left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}\left(\frac{\partial{P}}{\partial{T}}\right)\right]
\end{eqnarray}
\end{footnotesize}