### 2019-04-11 paper/derivatives.tex
Updates to derivative appendix, prepping for param. study

Derivatives of pressure with respect to $\tau$, $\epsilon$, and $D_{e}$. \\
\begin{footnotesize}
\begin{eqnarray}
	\left(\frac{\partial{P}}{\partial{\epsilon}}\right)_{\tau,D_{e}} &=& \left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}_{\rho, y_{e}}\left(\frac{\partial{P}}{\partial{T}}\right)_{\rho, y_{e}} \\
	\left(\frac{\partial{P}}{\partial{D_{e}}}\right)_{\tau, \epsilon} &=& \tau \left[
					\left(\frac{\partial{P}}{\partial{y_e}}\right)_{\rho, T} -
          \left(\frac{\partial{\epsilon}}{\partial{y_e}}\right)_{\rho, T}
					\left(\frac{\partial{P}}{\partial{\epsilon}}\right)_{\tau, D_{e}}\right]\\
	\left(\frac{\partial{P}}{\partial{\tau}}\right)_{\epsilon, D_{e}} &=& -\tau^{-2}
					\left[ \left(\frac{\partial{P}}{\partial{D_{e}}}\right)_{\tau, \epsilon}
					(y_e - 1) + \left( \frac{\partial{\epsilon}}{\partial{\rho}} \right)_{y_{e}, T}
          \left(\frac{\partial{P}}{\partial{\epsilon}}\right)_{\tau, D_{e}}\right]
\end{eqnarray}
\end{footnotesize}