### 2020-07-10 paper/derivatives.tex
fixed some bugs

We have an equation of state $P \,=\, P(\tau, \epsilon, D_{e})$ with $\tau \,=\, \rho^{-1}$, 
$\epsilon \,=\, e/ \rho$, and $D_{e} = \rho \, Y_{e}$. On the other hand, 
our tabulated equation of state constructs thermodynamic variables in terms of $\rho$, $T$, and $Y_{e}$. We 
need to express the thermodynamic derivatives of pressure necessary for characteristic limiting in terms of the table variables.
We start with the differential of pressure

\begin{equation}
 \label{eq:pressure_var1}
  dP = (\partial_{\tau}P)_{\epsilon, D_{e}}d\tau + (\partial_{\epsilon}P)_{\tau, D_{e}}d\epsilon + (\partial_{D_{e}}P)_{\epsilon, \tau}dD_{e}.
\end{equation}

Similarly, we may express the differentials of $\tau$, $\epsilon$, and $D_{e}$ in terms of the table variables

\begin{equation}
\begin{split}
  d\tau &= - \rho^{-2} d\rho\\
  d\epsilon &= (\partial_{\rho}\epsilon)_{T,Y_{e}}d\rho + (\partial_{T}\epsilon)_{\rho,Y_{e}}dT + (\partial_{Y_{e}}\epsilon)_{T,\rho}dY_{e}\\
  dD_{e} &= (\partial_{\rho}D_{e})_{T,Y_{e}}d\rho + (\partial_{T}D_{e})_{\epsilon,Y_{e}}dT + (\partial_{Y_{e}}D_{e})_{T,\epsilon}dY_{e}\\
   & = Y_{e} d\rho + \rho dY_{e}
\end{split}
\end{equation}

Inserting these differentials into Equation~\ref{eq:pressure_var1}, we find an expression for the pressure differential

\begin{equation} \label{eq:pressure_dif}
\begin{split}
  dP = &\left[ -\rho^{2} (\partial_{\tau}P)_{\epsilon, D_{e}} + (\partial_{\rho}\epsilon)_{T,Y_{e}}\,(\partial_{\epsilon}P)_{\tau, D_{e}}
    + Y_{e} (\partial_{D_{e}}P)_{\epsilon, \tau} \right] d\rho \\
    & + (\partial_{T}\epsilon)_{\rho,Y_{e}} (\partial_{\epsilon}P)_{\tau, D_{e}} dT \\
	& + \left[ (\partial_{Y_{e}}\epsilon)_{T,\rho} (\partial_{\epsilon}P)_{\tau, D_{e}} + \rho (\partial_{D_{e}}P)_{\epsilon, \tau}\right]dY_{e}   
\end{split}
\end{equation}

On the other hand, we have the differential of pressure in terms of the table variables

\begin{equation}
 \label{eq:pressure_table}
  dP = (\partial_{\rho}P)_{T, Y_{e}}d\rho + (\partial_{T}P)_{\rho, Y_{e}}dT + (\partial_{Y_{e}}P)_{T, \rho}dY_{e}.
\end{equation}

Comparing Equation~\ref{eq:pressure_dif} and Equation~\ref{eq:pressure_table}, we have the system of equations

\begin{gather}
  \begin{bmatrix} 
 	-\rho^{2} & (\partial_{\rho}\epsilon)_{T,Y_{e}} & Y_{e} \\ 
 	0 & (\partial_{T}\epsilon)_{\rho,Y_{e}} & 0 \\
	 0 & (\partial_{Y_{e}}\epsilon)_{T,\rho} & \rho
  \end{bmatrix}
  \begin{bmatrix}
	(\partial_{\tau}P)_{\epsilon, D_{e}} \\ (\partial_{\epsilon}P)_{\tau, D_{e}} \\ (\partial_{D_{e}}P)_{\epsilon, \tau}
  \end{bmatrix}
 =
  \begin{bmatrix}
	(\partial_{\rho}P)_{T, Y_{e}} \\ (\partial_{T}P)_{\rho, Y_{e}} \\ (\partial_{Y_{e}}P)_{T, \rho}
  \end{bmatrix}
\end{gather}

Solving, with some simplifications, we find the derivatives of pressure with respect to $\tau$, $\epsilon$, and $D_{e}$
in terms of the table variables $\rho$, $T$, $Y_{e}$. \\

\begin{footnotesize}
\begin{eqnarray}
	\left(\frac{\partial{P}}{\partial{\epsilon}}\right)_{\tau,D_{e}} &=& \left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}_{\rho, Y_{e}}\left(\frac{\partial{P}}{\partial{T}}\right)_{\rho, Y_{e}} \\
	\left(\frac{\partial{P}}{\partial{D_{e}}}\right)_{\tau, \epsilon} &=& \tau \left[
					\left(\frac{\partial{P}}{\partial{Y_{e}}}\right)_{\rho, T} -
          \left(\frac{\partial{\epsilon}}{\partial{Y_{e}}}\right)_{\rho, T}
					\left(\frac{\partial{P}}{\partial{\epsilon}}\right)_{\tau, D_{e}}\right]\\
	\left(\frac{\partial{P}}{\partial{\tau}}\right)_{\epsilon, D_{e}} &=& \tau^{-2}
					\left[ Y_{e}\left(\frac{\partial{P}}{\partial{D_{e}}}\right)_{\tau, \epsilon}
					 + \left( \frac{\partial{\epsilon}}{\partial{\rho}} \right)_{Y_{e}, T}
          \left(\frac{\partial{P}}{\partial{\epsilon}}\right)_{\tau, D_{e}} - 
		  \left(\frac{\partial{P}}{\partial{\rho}}\right)_{T, Y_{e}} \right]
\end{eqnarray}
\end{footnotesize}